Sampling Strategies in Decisions from Experience

### 2021-10-02 manuscript.Rmd
Review of Simulation Analysis

@@

cpt_agent %>% 
  ggplot(aes(x = s, y = gamma)) +
  geom_jitter(color = "#CECECE" ) + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Curvature ", gamma)),
       x ="Switching Probability",
       y = expression(gamma), 
       color="Switching Probability") + 
  theme_minimal()

@@

cpt_agent %>% 
  ggplot(aes(x = s, y = delta)) +
  geom_jitter(color = "#CECECE") + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 3) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 10, .5)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Elevation ", delta)),
       x ="Switching Probability",
       y = expression(delta), 
       color="Switching Probability") + 
  theme_minimal()

# weighting function

@@

## plot curves

cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

@@

@@

>>>>

The figures below display the estimates of the $\gamma$ and $\delta$ parameter of the probability weighting function [@prelecProbabilityWeightingFunction1998] fitted to DfE simulated for different parameter values of the generating model.
The estimates are plotted against the switching probability, where each panel represents a distinct boundary type and level, i.e., number of comparisons. 
Grey dots represent agent level estimates, colored dots represent the mean across all agent level estimates.

Most significantly, there is a strong relationship between the switching probability in the generating model and the $\gamma$ parameter.
I.e., large switching probabilities, which are indicative for small sample sequences, lead to larger estimates for $\gamma$. 
The resulting strong curvature leads to a compression of probabilities in the lower and upper range, reflecting underweighting of small probabilities and overweighting of large probabilities. 
This pattern is robust for varying degrees of the boundary level, except for $a = 1$. 
The latter deviation may be explained by the potential scale reduction factors for the respective parameter estimates, which indicate that the MCMC chains did not converge (see [Appendix 3][Appendix 3: Convergence Diagnostics for CPT Parameters]).

@@

cpt_agent %>% 
  ggplot(aes(x = s, y = gamma)) +
  geom_jitter(color = "#CECECE" ) + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 2) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Curvature ", gamma)),
       x ="Switching Probability",
       y = expression(gamma), 
       color="Switching Probability") + 
  theme_minimal()

@@

cpt_agent %>% 
  ggplot(aes(x = s, y = delta)) +
  geom_jitter(color = "#CECECE") + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 2) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 10, .5)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(paste("Elevation ", delta)),
       x ="Switching Probability",
       y = expression(delta), 
       color="Switching Probability") + 
  theme_minimal()
```

Below, the resulting probability weighting functions are displayed.

```{r}
# weighting function

@@

## plot curves
cpt_agent_w %>% 
  ggplot(aes(x = ep, y = w)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_w, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) + 
  labs(title = "Probability Weighting Function",
       x ="Experienced Probability",
       y = "Subjective Experienced Probability", 
       color="Switching Probability") + 
  theme_minimal()

@@

The false response rates for different parameter values of the generating model reflect the probability weighting patterns from above. 
That is, the strong curvature resulting from large switching probabilities produces an underweighting of small probabilities. This in turn has the effect that the rarity of an attractive (unattractive) outcome leads to higher rates of choosing the safe (risky) prospect although the risky (safe) prospect had a higher experienced expected value.

# compute false response rates

fr_rates <- choices %>% 
  mutate(ev_ratio_exp = round(a_ev_exp/b_ev_exp, 2), # experienced EV (eEV)
         norm = case_when(ev_ratio_exp > 1 ~ "A", ev_ratio_exp < 1 ~ "B")) %>% # normative choice according to eEV
  filter(!is.na(norm)) %>% # exclude trials with normative indifferent prospects
  group_by(s, boundary, a, rare, norm, choice) %>% # group correct and incorrect responses
  summarise(n = n()) %>% # absolute numbers 
  mutate(rate = round(n/sum(n), 2), # response rates 
         type = case_when(norm == "A" & choice == "B" ~ "false safe", norm == "B" & choice == "A" ~ "false risky")) %>% filter(!is.na(type))  # remove correct responses

fr_rates %>% 
  ggplot(aes(a, s, fill = rate)) + 
  geom_tile(colour="white", size=0.1) +
  scale_fill_scico(palette = "buda") + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  scale_x_continuous(expand=c(0,0), breaks = seq(1, 5, 1)) +
  scale_y_continuous(expand=c(0,0), breaks = seq(.1, 1, .1)) +
  labs(x = "a", 
       y= "s", 
       fill = "% False Responses") + 
  theme_minimal()

fr_rates %>% 
  ggplot(aes(s, rate, color = s)) + 
  facet_grid(type ~ fct_relevel(rare, "attractive", "none", "unattractive"), switch = "y") +
  geom_jitter(size = 3) + 
  scale_x_continuous(breaks = seq(0, 1, .1)) +
  scale_y_continuous(breaks = seq(0, 1, .1)) +
  scale_color_scico(palette = "buda") + 
  labs(title = "Piecewise Integration",
       x = "s", 
       y= "% False Responses", 
       color = "a") + 
  theme_minimal()
```

@@

The figures below display the estimates of the $\alpha$ and the resulting value function.

# alpha
cpt_agent %>% 
  ggplot(aes(x = s, y = alpha)) +
  geom_jitter(color = "#CECECE" ) + 
  geom_point(data = cpt_agent_mean, aes(color = s), size = 2) +
  geom_line(data = cpt_agent_mean, aes(color = s), size = 1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 1, .2)) + 
  scale_y_continuous(breaks = seq(0, 2, .2)) +
  facet_grid(boundary~a, switch = "y") +
  labs(title = expression(alpha),
       x ="Switching Probability",
       y = expression(alpha), 
       color="Switching Probability") + 
  theme_minimal()

# value function

## compute values

cpt_agent_v <- cpt_agent %>% 
  select(-c(gamma, delta, rho)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

cpt_agent_mean_v <- cpt_agent %>% 
  select(-c(gamma, delta, rho)) %>% 
  group_by(s) %>% 
  summarise(alpha = round(mean(alpha), 2)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

## plot curves
cpt_agent_v %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(color = "#CECECE") + 
  geom_path(data = cpt_agent_mean_v, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()

cpt_agent_mean_v <- cpt_agent %>% 
  group_by(s, a) %>% 
  summarise(alpha = round(mean(alpha), 2)) %>% 
  expand_grid(x = seq(0, 20, 2)) %>%  
  mutate(v = round(x^alpha, 2))

cpt_agent_v %>% 
  ggplot(aes(x = x, y = v)) +
  geom_path(color = "#CECECE", size = 0.5) + 
  geom_path(data = cpt_agent_mean_v, aes(color = s), size = 1.1) +
  scale_color_scico(palette = "buda") + 
  scale_x_continuous(breaks = seq(0, 20, 5)) +
  scale_y_continuous(breaks = seq(0, 20, 5)) + 
  facet_wrap(~a, nrow = 2) + 
  labs(title = "Value Function",
       x ="Objective Value",
       y = "Subjective Value", 
       color="Switching Probability") + 
  theme_minimal()
```
